### 2012-05-17 lucene/handout.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{fontspec}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{parskip}
\usepackage{sectsty}
\usepackage{unicode-math}

\defaultfontfeatures{Scale=MatchLowercase, Mapping=tex-text}
\setmainfont{Optima LT Std}
\setmathfont{Neo Euler}

\sectionfont{\fontspec[Scale=1.44]{Optima LT Std}}
\subsectionfont{\fontspec[Scale=1.2]{Optima LT Std}}

\begin{document}
\pagestyle{empty}
\Large

\section*{Indexing documents}

\begin{enumerate}
\item Parse XML: extract title, subtitle, abstract, full text, metadata
\item Tokenize, normalize, filter stopwords
\item Build inverted index
\end{enumerate}

\section*{Searching documents}

\subsection*{Filter documents using the Boolean Model}

\begin{enumerate}
\item Tokenize, normalize, filter the terms of the query
\item Rewrite the query in disjunctive normal form
\item Return documents which satisfy the query
\end{enumerate}

\subsection*{Score documents using the Vector Space Model}

Document and query are high-dimensional vectors

Cosine similarity:

\begin{equation*}
\mathrm{similarity}(q, d) = \frac{V(q) \cdot V(d)}{\vert V(q) \vert \vert V(d)
  \vert}
\end{equation*}

Adapted for Lucene:

\begin{equation*}
\mathrm{score}(q, d) = \mathrm{coord}(q,d) \mathrm{queryBoost}(q) \frac{V(q)
  \cdot V(d)}{\vert V(q) \vert} \mathrm{docNorm}(d) \mathrm{docBoost}(d)
\end{equation*}

Precompute some things and further generalize:

\begin{equation*}
  \mathrm{score}(q,d) = \mathrm{coord}(q,d)\mathrm{queryNorm}(q) \sum_{t \in q}
  \left( \mathrm{tf}(t, d) \mathrm{idf}(t)^2 \mathrm{boost}(t) \mathrm{norm}(t,
  d) \right)
\end{equation*}

\end{document}